# Reproducible Research: Peer Assessment 1

## Loading packages
```{r}
library(plyr)
```
## Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv")
str(activity)
summary(activity)
```

Create new data frame for daily activity with two variables, dates and spd (steps per day):

```{r}
daily <- ddply(activity, .(date), summarise, steps = sum(steps,na.rm=TRUE))
activity$interval <- as.factor(activity$interval)
per_min <- ddply(activity, .(interval), summarise, steps = sum(steps,na.rm=TRUE))
str(per_min)
```

## What is mean total number of steps taken per day?

```{r}
 hist(daily$steps,nrow(daily))
```

The mean total number of steps taken per day is `r mean(daily$steps, na.rm=TRUE)` steps. The median is `r median(daily$steps,na.rm=TRUE)` steps.

## What is the average daily activity pattern?

```{r}
plot(per_min$interval,per_min$steps)
```

The 5-minute interval when most steps are taken (on average) is the `r per_min$interval[per_min$steps == max(per_min$steps)]`th interval.

## Imputing missing values


The total number of missing values is `r sum(is.na(activity))`.

## Are there differences in activity patterns between weekdays and weekends?
